---
title: "Exam work"
author: "EJ Arce"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library(stats)
```

### 2Ia

(.9^15) + (15*.1*(.9^14)) + (15*7*(.1^2)*(.9^13))

pbinom(2, size = 15, p = .1)

###2IB


```{r}
((choose(17, 13)*choose(3,2)) + (choose(17, 14)*choose(3,1)) + (choose(17, 15)))/choose(20, 15)
```

###3D
```{r}
choose(10, 2)*choose(8, 5)/(3^10)
```

###4A

```{r}
(choose(4,3)*choose(4,2)*choose(44,8))/(choose(52,13))
```

###4B

```{r}
(choose(4,2)*choose(4,1)*choose(44,10))/choose(52,13)
```


###4C

```{r}
choose(35, 13) / choose(39, 13)
```

###4D

```{r}
(choose(17,3)*choose(3,1) + choose(14,1))/choose(20,5)
```


###2 Part II

```{r}
set.seed(30)
worms <- c(0:15)
nsim = 10000
probs <- c()
count = 0
for(i in 0:15){
  for(j in 1:nsim){
    apples <- sample(0:1, size = 15, replace = T, prob = c(.9, .1))
      if(i %in% sum(apples)){
        count = count+1
      }
  }
  probs <- c(probs, count/nsim)
}
cbind(worms, probs)
```

The simulation returns a vector with 16 rows and 2 columns. The worms column indicates the simulated probability of obtaining, at most, that many apples with worms. The next vector is the simulated probabilties of picking these cumulative numbers of apples containing worms. So the probability of obtaining at most 2 apples with worms is about .8072.

### 7C

```{r}
z <- rbinom(size = 5, n = 1000, p =.5)
hist(z)
```

