---
title: "Problem Set 2 R exercises"
author: "EJ Arce"
date: "9/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("boot")
library("dplyr")
```

### 2.28a

```{r}
# I will set a seed and run 10000 simulations. We will calculate the number of
# simulations with a match (count), and keep updating that value after the each
# simulation. Then count/nsim will give us our simulated probability.
set.seed(30)
nsim = 10000
count = 0

for(i in 1:nsim){
  birthdays <- sample(1:365, 23, replace = T)
   if(2 %in% table(birthdays)){
    count = count+1
  }
}
count/nsim
```

In a room of 23 people, there is almost a 50% chance that two people will have the same birthday.

###2.28b

```{r}
# We will solve for the size of the room by repeating the loop process for
# different sizes, in increments of 10, ranging from 10 to 50. A vector with
# simulated probabilities for a given size will be returned.
set.seed(30)
size <- seq(10, 50, by = 5)
nsim = 1000
count = 0
probs <- c()

for(i in 1:length(size)){
  for(j in 1:nsim){
    birthdays <- sample(1:365, size[i], replace = T)
      if(2 %in% table(birthdays)){
        count = count+1
      }
  }
  probs <- c(probs, count/nsim)
}
cbind(size, probs)
```

We see that between 45 and 50 people are needed. To find an exact number, we will repeat the process for only sizes 45-50.

```{r}
# Repeating the same command, only specifying the size sequence.
set.seed(23)
size <- seq(45, 50, by = 1)
nsim = 10000
probs <- c()

for(i in 1:length(size)){
  count = 0
  for(j in 1:nsim){
    birthdays <- sample(1:365, size[i], replace = T)
      if(2 %in% table(birthdays)){
        count = count+1
      }
  }
  probs <- c(probs, count/nsim)
}
cbind(size, probs)
```

A room of about 48 people will have a 95% chance that two people will have the same birthday.

### 2.28c

```{r}
# Repeating the command for 2.28a, changing the number of matches to 3, and the
# number of people in a room to 50.
set.seed(11)
nsim = 10000
count = 0

for(i in 1:nsim){
  birthdays <- sample(1:365, 50, replace = T)
   if(3 %in% table(birthdays)){
    count = count+1
  }
}
count/nsim
```

In a room of 50 people, there is about a 12% chance that three people will have the same birthday.

### 2.28d

```{r}
# Repeating the command for 2.28b, changing the number of matches to 3, and
# solving the room size for a probability of .50.
set.seed(35)
size <- seq(10, 100, by = 10)
nsim = 10000
probs <- c()

for(i in 1:length(size)){
  count = 0
  for(j in 1:nsim){
    birthdays <- sample(1:365, size[i], replace = T)
      if(3 %in% table(birthdays)){
        count = count+1
      }
  }
  probs <- c(probs, count/nsim)
}
cbind(size, probs)
```

Similar to b, we will repeat the process for sizes 80-90 only.

```{r}
# Running the same command with the sequance being integers from 80 to 90.
set.seed(35)
size <- seq(80, 90, by = 1)
nsim = 10000
probs <- c()

for(i in 1:length(size)){
  count = 0
  for(j in 1:nsim){
    birthdays <- sample(1:365, size[i], replace = T)
      if(3 %in% table(birthdays)){
        count = count+1
      }
  }
  probs <- c(probs, count/nsim)
}
cbind(size, probs)
```

About 89 people need to be in a room together so that the probability of three people having the same birthday is 50%.
