---
title: "Late game situations"
author: "EJ Arce"
date: "11/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Late game probability situations
#### 1

Suppose each player is tied and has the following deck

* 8 copper
* 3 silver
* 3 gold
* 2 cellar
* 2 village
* 2 throne rooms
* 2 witch

Suppose there is one province left. Find the probability that a player's hand has enough money to buy the last province and win the game.

```{r}
# Calculate sample space
ss <- choose(22,5)
# Calculate combinations of hands with at least 8 dollars

# 3 golds and 2 non-golds
gold3 <- choose(3,3)*choose(19,2)

# 2 golds and either: >= 1 silver, 2 non-gold or 0 silvers, >= 2 coppers
gold2 <- choose(3,2)*(choose(3,1)*choose(18,2) + choose(8,2)*choose(14,1))

# 1 gold and either: 1 silver, 3 copper or 2 silver, 1 copper, 1 non-gold
gold1 <- choose(3,1)*(choose(3,1)*choose(8,3) +
                        choose(3,2)*choose(8,1)*choose(16,1))

# 0 gold, 3 silvers, 2 coppers
gold0 <- choose(3,3)*choose(8,2)

p.province <- (gold3 + gold2 + gold1 + gold0)/ss
p.province
```

NOTE: THIS DOES NOT CONSIDER THE CHANCE OF GAINING MORE CARDS TO DRAW

### 2

Suppose the following:

* game there is one province left and everyone has the same deck as above
* you are ordered third and have 5 less estate than the player in first place
* you have enough money to buy the last province

What is the probability you get cursed and cannot win the game?

notes: you can get cursed, but you can prevent the curse with a moat