---
title: "Late game situations"
author: "EJ Arce"
date: "11/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Late game probability situations

For the following problems, suppose that each player's deck must be re-shuffled after every turn.

#### 1

Suppose each player is tied and has the following deck:

* 8 copper
* 3 silver
* 3 gold
* 3 duchies
* 3 provinces

Suppose there is one province left. Find the probability that a player's hand has enough money to buy the last province and win the game.


```{r}
# Calculate sample space
ss.1 <- choose(22,5)
# Calculate combinations of hands with at least 8 dollars

# 3 golds and 2 non-golds
gold3 <- choose(3,3)*choose(19,2)

# 2 golds and either: >= 1 silver, 2 non-gold or 0 silvers, >= 2 coppers
gold2 <- choose(3,2)*(choose(3,1)*choose(18,2) + choose(8,2)*choose(14,1))

# 1 gold and either: 1 silver, 3 copper or 2 silver, 1 copper, 1 non-gold
gold1 <- choose(3,1)*(choose(3,1)*choose(8,3) +
                        choose(3,2)*choose(8,1)*choose(16,1))

# 0 gold, 3 silvers, 2 coppers
gold0 <- choose(3,3)*choose(8,2)

p.province <- (gold3 + gold2 + gold1 + gold0)/ss.1
p.province
```

##### Simulating 1

```{r}
# Need a for loop returning 5 cards of random digits indicating money value
# 8 0's, 8 1's, 3 2's, 3 3's

l <- rbind(c(0,0,0,0,0,0,0,0,
            1,1,1,1,1,1,1,1,
            2,2,2,3,3,3))
set.seed(30)
count <- 0

for(i in 1:10000){
  hands <- sample(l, 5, replace = F)
  if(sum(hands) >= 8){
    count = count + 1
  }
}
count/10000
```


#### 2

Assume the game will only last for 20 more rounds and that buying is not permitted during this period. Suppose each player has the following cards:

* 8 copper
* 1 silver
* 4 gold
* 7 estates
* 2 duchies
* 3 provinces
* 5 witches
* 1 moat
* 2 festivals

In this version, players will reshuffle the deck after each of their turns. If a witch is drawn, every player wants to play it as their action.

How many curses do you expect to counter with a moat?

To counter a witch with a moat in a given round, an opponent must play a witch while your hand contains a moat.

```{r}
# A player can have 5 cards if the moat is not picked or 7 cards if the moat is
# picked. There are 33 cards in your deck.
ss.2 <- choose(32,5) + choose(1,1)*choose(32,6)

# A witch drawn in a hand of 5 cards without a moat or 7 cards with a moat by 
# either opponent.
p.witch <- 2*(choose(5,1)*choose(32,4) +
                choose(1,1)*choose(31,5)*choose(5,1)) / ss.2

# A moat must be contained in a hand of 5
p.moat <- choose(1,1)*choose(33,4)/choose(33,5)

# Probability of drawing a moat and either opponent drawing a witch.
p.counter <- p.moat*p.witch
p.counter
```

The probability of an opponent playing a curse and you being able to counter with a moat is .00531. The number of these counters in 20 rounds (40 opponent turns) follows a binomial distribution with n = 40, and p = .00531. Thus the expected number of counters is np

```{r}
p.counter*40
```

You should expect to counter 3.63 curse plays.

##### Simulation

* 8 copper
* 1 silver
* 4 gold
* 7 estates
* 2 duchies
* 3 provinces
* 5 witches
* 1 moat
* 2 festivals

```{r}
library(dplyr)
card <- cbind(c(0,0,0,0,0,0,0,0,0,0,0,0,
                0,0,0,0,0,0,0,0,0,0,0,0,
                0,0,0,1,2,2,2,2,2))
cards <- data.frame(card)

sims <- c()
nsim <- 1000

for(i in 1:nsim){
  count = 0
  for(j in 1:40){
    opphands <- sample(1:nrow(cards),5,replace = F)
    deck <- data.frame(cards[-opphands,])
    opphands <- data.frame(cards[opphands,])
    opphands <- rename(opphands, card = cards.opphands...)
    if(1 %in% opphands$card){
      plus2 <- sample(1:nrow(deck), 2, replace = F)
      plus2 <- data.frame(deck[plus2,])
      plus2 <- rename(plus2, card = deck.plus2...)
      opphands <- rbind(plus2,opphands)
    }
  if(2 %in%  opphands$card){
    myhand <- sample(1:nrow(cards), 5, replace = F)
    myhand <- data.frame(cards[myhand,])
    if(1 %in% myhand$cards.myhand...){
      count = count + 1
      }
    }
  }
  sims <- c(sims, count)
}
  
mean(sims)
```

